Can Cultural Transmission Explain The Evolution of Cooperative Behavior?

### 2020-07-04 ms/main.tex
fix merge issue
Add more results - find stability of equilibrium points

\textbf{Result 2.1: equliburims with equal horizontal transmission $T$ and vertical transmission $0<v<1$}
\begin{equation} \label{eq:result2.1}
  \hat{p}^*_0 =  0, \, \hat{p}^*_1 = 1
\end{equation}
We would like to know under what condition fixation of cooperation is stable. 
In order to do that we find the derivative  $f(\hat{p}) = \hat{p}'-\hat{p}$.
\begin{equation} \label{eq:derivative_of_phattag-phat}
 f'(\hat{p}) = 2\cdot\hat{p} (-\alpha bvT + cv(1-T)) + (\alpha bvT -cv(1-T)).
\end{equation}
We are interested in the case when when $\hat{p} = \hat{p}^*_1=1$ therefore
$f'(1) = -\alpha bvT + cv(1-T)$. Therefore, fixation of cooperation is stable when
\textbf{Results 2.2: condition for stability of fixation of cooperation}
\begin{equation} \label{eq:oblique_and_vertic_result}
  \frac{b}{c}>\frac{1-T}{\alpha T}
\end{equation}
Which is the same condition we get in corollary 1.1.
Now let's assume that the general case where $T_A \neq T_B$ and we get that
\begin{equation} \label{eq:general_case_polynomial}
  \hat{p}'-\hat{p} = \beta_1p^3 + \beta_2 p^2 + \beta_3p + \beta_4
\end{equation}
where 
\begin{equation} \label{eq:polynomial_coefficients}
\begin{split}
  & \beta_1 = c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B) \\
  & \beta_2 = (b(1-\alpha v)-c(1-v)-1)(T_A-T_B)-\alpha bv T_A + cv(1-T_B) \\
  & \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) \\
  & \beta_4 = 0
\end{split}
\end{equation}
Since the degree of eq.~\ref{eq:general_case_polynomial} is a third degree polynomial,
three equilibrium exist. We already know that $\hat{p}^*_0 = 0 $ and $\hat{p}^*_1 = 1$ are equilibrium.
We can find the third equilibrium by finding wehen eq.~\ref{eq:general_case_polynomial} is equal to zero and we get
\begin{equation} \label{eq:oblique_and_vertic_result}
  \hat{p}^*_2 =  \frac{\alpha bvT_A + cvT_B + T_A-T_B-cv}{c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B)}
\end{equation}
Next we would like to learn under what conditions $\hat{p}^*_2$ is locally stable. 
For simplicity let's assume that $T_A > T_B$ and we will check when $\hat{p}^*_0 \, \hat{p}^*_1$ are stable.
In order to do that we find the derivative  $f(\hat{p}) = \hat{p}'-\hat{p}$ of $\hat{p}$.
\begin{equation} \label{eq:derivative_of_phattag-phat2}
  f'(\hat{p}) = 3\beta_0 \hat{p}^2 + 2\beta_1\hat{p}^2 + \beta_3
 \end{equation}
 When $\hat{p} = \hat{p}^*_0 = 0$ we get
 \begin{equation} \label{eq:derivative_of_phattag-phat_p=0}
  f'(0) = \beta_3
 \end{equation}
 Therefore, $\hat{p}^*_0$ is stable when
 \begin{equation} \label{eq:p=0_stability_condition}
  f'(0) = \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) < 0
 \end{equation}
 And $\hat{p}^*_1=1$ is stable when
 \begin{equation} \label{eq:p=1_stability_condition}
  f'(1) =3\beta_1 + 2\beta_2 +\beta_3 = cv + (T_A-T_B)(b-c) - T_B(1+\alpha bv)  + T_A(1-cv) < 0
 \end{equation}

>>>>>>> Stashed changes
\begin{figure}[H]
  \centering
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2a.pdf}
    \caption{$V=1$, $T_A=T_B=T$, $\alpha \neq 0$}
    \label{fig:results_a}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2b.pdf}
    \caption{$V=1$, $\alpha = 0$}
    \label{fig:results_b}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2c.pdf}
    \caption{$V=0$}
    \label{fig:results_c}
  \end{subfigure}
  \label{fig:results}
  \caption{numerical results}
\end{figure}

>>>>

\textbf{Result 2.1: equliburims with equal horizontal transmission $T$ and vertical transmission $0<v<1$}
\begin{equation} \label{eq:result2.1}
  \hat{p}^*_0 =  0, \, \hat{p}^*_1 = 1
\end{equation}
We would like to know under what condition fixation of cooperation is stable. 
In order to do that we find the derivative  $f(\hat{p}) = \hat{p}'-\hat{p}$.
\begin{equation} \label{eq:derivative_of_phattag-phat}
 f'(\hat{p}) = 2\cdot\hat{p} (-\alpha bvT + cv(1-T)) + (\alpha bvT -cv(1-T)).
\end{equation}
We are interested in the case when when $\hat{p} = \hat{p}^*_1=1$ therefore
$f'(1) = -\alpha bvT + cv(1-T)$. Therefore, fixation of cooperation is stable when
\textbf{Results 2.2: condition for stability of fixation of cooperation}
\begin{equation} \label{eq:oblique_and_vertic_result2}
  \frac{b}{c}>\frac{1-T}{\alpha T}
\end{equation}
Which is the same condition we get in corollary 1.1.
Now let's assume that the general case where $T_A \neq T_B$ and we get that
\begin{equation} \label{eq:general_case_polynomial}
  \hat{p}'-\hat{p} = \beta_1p^3 + \beta_2 p^2 + \beta_3p + \beta_4
\end{equation}
where 
\begin{equation} \label{eq:polynomial_coefficients}
\begin{split}
  & \beta_1 = c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B) \\
  & \beta_2 = (b(1-\alpha v)-c(1-v)-1)(T_A-T_B)-\alpha bv T_A + cv(1-T_B) \\
  & \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) \\
  & \beta_4 = 0
\end{split}
\end{equation}
Since the degree of eq.~\ref{eq:general_case_polynomial} is a third degree polynomial,
three equilibrium exist. We already know that $\hat{p}^*_0 = 0 $ and $\hat{p}^*_1 = 1$ are equilibrium.
We can find the third equilibrium by finding when eq.~\ref{eq:general_case_polynomial} is equal to zero and we get
\begin{equation} \label{eq:oblique_and_vertic_result}
  \hat{p}^*_2 =  \frac{\alpha bvT_A + cvT_B + T_A-T_B-cv}{c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B)}
\end{equation}
We would like to know when $\hat{p}_2$ is realistic meaning $0<\hat{p}_2<1$.  
In order to do that we find the derivative  $f(\hat{p}) = \hat{p}'-\hat{p}$ of $\hat{p}$.
\begin{equation} \label{eq:derivative_of_phattag-phat2}
  f'(\hat{p}) = 3\beta_0 \hat{p}^2 + 2\beta_1\hat{p}^2 + \beta_3
 \end{equation}
We assume that $T_A>T_B$ and we get that $\beta_1<0$ since $c(1-v)-b(1-\alpha v)<0$ and $c<b$ $v<1$.
In this case the polynomial goes to $\infty$ when $\hat{p}$ goes to infinity. 
Therefore, if $\hat{p}_2$ is positive then $f'(0)$ must be negative
\begin{equation} \label{eq:derivative_of_phattag-phat_p=0}
  f'(0) = \beta_3
 \end{equation}
$\hat{p}_2>0$ when $\beta_3<0$. In addition, $\hat{p}_0=0$ is stable when $|f'(0)|<1$ therefore, when $-1<\beta_3<1$. We can prove that $-1<\beta_3$:
\begin{equation} \label{eq:beta3-condition}
  \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) > -cv(1-T_B) > -1
\end{equation}
Therefore, when $\hat{p}_2>0$ we get that $-1<\beta_3<0$ and $\hat{p}_0=0$ is always stable.

Now we would like to find when $\hat{p}_2<1$. If $T_A>T_B$ like before then the polynomial goes to minus infinity when $\hat{p}$ goes to infinity.
Hence, if $\hat{p}_2<1$ then $f'(1)$ must be negative
\begin{equation} \label{eq:derivative_of_phattag-phat_p=1}
  f'(0) = \beta_1 - \beta_3
\end{equation}
Therefore, $\hat{p}_2<1$ when $\beta_1-\beta_3<0$. In addition, $\hat{p}_1=1$ is stable when $|f'(1)|<1$ therefore, when $-1<\beta_1-\beta_3<1$. 
We can prove that when $0<\hat{p}_2<1$ hence $\beta_3<0$ and $\beta_1-\beta_3<0$ we get that $-1<\beta_1-\beta_3$:
\begin{equation} \label{eq:beta1-beta3-condition}
 \beta_1 - \beta_3 > - \beta_3 > 0 > -1
\end{equation}
which means that when $0<\hat{p}_2<1$ then $\hat{p}_1=1$ is always stable.

\textbf{result 2.3}: for the case when $T_A>T_B$ $\hat{p_2}$ is realistic when $\beta_1<\beta_3<0$ and is always unstable equilibrium.

\textbf{result 2.4}: for the case when $T_B>T_A$ $\hat{p_2}$ is realistic when $\beta_1>\beta_3>0$ and is always unstable equilibrium.

\begin{figure}[H]
  \centering
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2a.pdf}
    \caption{$V=1$, $T_A=T_B=T$, $\alpha \neq 0$}
    \label{fig:results_a}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2b.pdf}
    \caption{$V=1$, $\alpha = 0$}
    \label{fig:results_b}
  \end{subfigure}
  \begin{subfigure}{8cm}
    \includegraphics[scale=0.5]{figure2c.pdf}
    \caption{$V=0$}
    \label{fig:results_c}
  \end{subfigure}
  \label{fig:results}
  \caption{numerical results}
\end{figure}
